---
title: "Introduction to Rthoptera"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Rthoptera}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Rthoptera)
```

---
title: "Introduction to Rthoptera"
author: "Francisco Rivas"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Rthoptera}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction to Rthoptera

**Rthoptera** is an R package that provides interactive Shiny applications for standard analysis of insect sounds. It offers tools for pre-processing, spectral and temporal analysis, as well as interactive visualization of oscillograms and spectrograms. This package aims to simplify the process of importing, processing, and analyzing insect sound recordings in a user-friendly interface using standardized plotting and measurements. 

## Pre-processing

The package offers several pre-processing tools that allow you to handle audio files before performing detailed analysis:

- **Import**: Browse your local data to import audio files (WAV) as Wave objects into R. During this process, all audio files are automatically corrected for DC offset to ensure accurate analysis.
  
- **Downsample**: When required, you can reduce the sampling rate of your audio files. This helps improve the computing speed for certain analyses and plots without losing important sound features.
  
- **High-pass Filter**: When necessary, apply a user-defined high-pass filter to eliminate low-frequency noise from your audio recordings. This is useful for removing unwanted background noise.
  
- **Trim**: Using an interactive oscillogram, you can visually select and trim specific sections of a Wave object. The trimmed sections can be saved as new Wave objects in your R environment for further analysis.

## Analysis

Once pre-processing is complete, Rthoptera offers a range of analysis tools to extract spectral and temporal statistics from your audio files:

- **Spectral Statistics**: Automatically calculate spectral metrics based on the mean power spectrum of a Wave object. These statistics provide insights into the frequency domain of the audio signal.

- **Temporal Statistics**: Automatically extract temporal metrics from your audio recordings. This includes identifying and analyzing peaks (pulses), trains (syllables), and echemes (groups of syllables or trills) in the insect sounds.

## Plotting

- **Multi-Power Spectra**: An interactive tool that overlays multiple power spectra on a color-coded oscillogram, allowing for easy comparison and visualization of spectral features across different time intervals.

- **Spectrogram**: Generate standard spectrograms alongside a lateral mean power spectrum. This combination allows you to visualize both time-frequency representations and the overall spectral distribution.

- **Multi Plot**: A combined visualization of the spectrogram and oscillogram, providing a comprehensive view of both the time-domain and frequency-domain characteristics of the insect sounds.

- **Oscillogram**: Create standard oscillograms as well as interactive oscillograms that allow you to zoom in and explore specific sections of the waveform.

- **Multi-oscillogram**: Create a stacked oscillogram plot for comparing the sounds of multiple species. This is particularly useful for analyzing the differences in acoustic patterns between species.

## Installation

To install the Rthoptera package, follow these steps:

1. Install the `remotes` package if you haven't already:

    ```r
    install.packages("remotes")
    ```

2. Load the `remotes` package:

    ```r
    library(remotes)
    ```

3. Install **Rthoptera** from GitHub:

    ```r
    remotes::install_github("panchorivasf/Rthoptera")
    ```

4. Load the **Rthoptera** package:

    ```r
    library(Rthoptera)
    ```

